### 2013-09-18 chi2014/related.tex

Following Osborne's work, Taylor et al.\ established that brainstorming by \emph{nominal groups}, or individuals working in isolation from one another, yields better results than collocated groups in terms of number of ideas generated \cite{taylor_does_1958}. These effects can be partially explained by the absence of counterproductive social forces, such as the fear of judgment. Bouchard and Hare \cite{bouchard_jr_size_1970}, later found that nominal brainstorming groups are able to generate quantities of ideas linear in the number of members in the group.

Electronic brainstorming is a variant in which individuals brainstorm at separate computer terminals, with the exchange of ideas performed over the network \cite{gallupe_electronic_1992}. Participants can enter ideas while simultaneously viewing the ideas of all other ideas, which are anonymously presented on-screen. As with nominal brainstorming, Gallupe \cite{gallupe_electronic_1992} observed that electronic brainstorming can reduce counterproductive social effects, such as production blocking (the inability for multiple people to speak at once) and fear of judgment, compared to physically collocated groups. However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, or individuals limiting their idea generation to avoid replicating the ideas of others. Given these drawbacks, their results suggest there is little evidence to support the proposition that electronic brainstorming outperforms nominal brainstorming.

@@

In terms of individual brainstorming processes, Nijstad and Stroebe provide a model for individual idea generation dubbed SIAM, or the search for ideas in associative memory. \cite{nijstad_how_2006}. SIAM defines two stages of idea generation: the generation of ideas based working memory, and idea production within a single activated ``image'' (where an image can be thought of as a category of ideas). This model predicts the following:

\begin{enumerate}
\item an idea is more likely to be followed by an idea in the same category than would be expected by chance
\item new ideas in the same category as the previous idea are generated faster than ideas between categories
\item increased clustering (generation of ideas within-category) is positively correlated with overall high production
\end{enumerate}

Later, we will present evidence in support of this model from the data we collected.

In the realm of crowd-based brainstorming, there are a number of ways this process could be realized. The most basic configuration mirrors that of nominal brainstorming, where individuals brainstorm ideas in isolation of one another, and the resulting ideas are passed on to the requestor (the individual posting the brainstorming task). More sophisticated brainstorming environments are possible, such as those that provide awareness of ideas that others have generated, or those that rely on the crowd to organize or rank the ideas produced by others. Examples for these environments are platforms for generating social innovation ideas or product concepts like OpenIDEO, Quirky and Threadless that employ structured process where participants take on different responsibility from generating ideas to evaluating others' ideas. In this paper, we focus on the most basic setup, where individuals brainstorm in isolation of one another.

While we are unaware of specific studies of brainstorming processes in crowd marketplaces, there is a range of related research examining more creative processes in this environment \cite{lewis2011affective, kittur2011crowdforge, Zhang:2012:HCT:2207676.2207708}. For example, Yu and Nickerson created a system in which individuals participate in an evolutionary algorithm to generate creative chair designs \cite{yu_cooks_2011}. Compared to ideas produced by single individuals, they found the ideas produced as a result of their genetic algorithm were more creative. However, individuals produced only one idea each, compared to the many ideas typically produced by individuals in a brainstorming session. Thus, while their approach demonstrates the creative potential of the crowd, their results say little about brainstorming processes by crowds.

>>>>

Following Osborne's work, Taylor et al.\ established that brainstorming by \emph{nominal groups}, or individuals working in isolation from one another whose results are later combined, yields better results than collocated groups in terms of number of ideas generated \cite{taylor_does_1958}. These effects can be partially explained by the absence of counterproductive social forces, such as production blocking (the inability for multiple people to speak at once) and the fear of judgment. Bouchard and Hare \cite{bouchard_jr_size_1970}, later found that nominal brainstorming groups are able to generate quantities of ideas linear in the number of members in the group.

Electronic brainstorming is a variant in which individuals brainstorm at separate computer terminals, with the exchange of ideas performed over the network \cite{gallupe_electronic_1992}. Participants can enter ideas while simultaneously viewing the ideas of all other ideas, which are anonymously presented on-screen. As with nominal brainstorming, Gallupe \cite{gallupe_electronic_1992} observed that electronic brainstorming can reduce counterproductive social effects. However, Pinsonneault et al.\ \cite{pinsonneault_electronic_1999} identify productivity impediments introduced by group electronic brainstorming, such as being distracted by other ideas appearing, or individuals witholding ideas they feel are derivative of others. Given these drawbacks, their results suggest there is little evidence to support the proposition that electronic brainstorming outperforms nominal brainstorming.

@@

In terms of individual brainstorming processes, Nijstad and Stroebe provide a model for individual idea generation dubbed SIAM, or the search for ideas in associative memory. \cite{nijstad_how_2006}. SIAM defines two stages of idea generation: the generation of ideas based working memory, and idea production within a single activated ``image'' (where an image can be thought of as a category of ideas). This model makes predictions applicable to all brainstorming contexts, namely that an idea is more likely to be followed by an idea in the same category than would be expected by chance, and new ideas in the same category as the previous idea are generated faster than ideas between categories. Nijstad and Stroebe verify these in the context of nominal brainstorming.

The simplist imaginable instantiation of brainstorming in a microtask marketplace mirrors that of nominal brainstorming. More sophisticated brainstorming environments are possible, such as those that provide awareness of ideas that others have generated (electronic brainstorming), or those that rely on the crowd to organize or rank the ideas produced by others, as in evolutionary techniques \cite{yu_cooks_2011}.
%Examples for these environments are platforms for generating social innovation ideas or product concepts like OpenIDEO, Quirky and Threadless that employ structured process where participants take on different responsibility from generating ideas to evaluating others' ideas.
In this paper, we examine nominal brainstorming to establish a baseline.

While we are unaware of specific studies of brainstorming in microtask marketplaces, there is a range of related research examining other creative processes \cite{lewis2011affective, kittur2011crowdforge, Zhang:2012:HCT:2207676.2207708}. For example, Yu and Nickerson created an evolutionary crowd algorithm to generate creative chair designs \cite{yu_cooks_2011}. Their work and others demonstrates the creative potential of the crowd, but does not specifically address brainstorming or analogous processes in which participants generate many ideas.

### 2013-09-18 chi2014/results.tex

Hierarchical clustering yielded an idea forest composed of 3007 instances, 1213 unique ideas, 335 category trees, and 129 non-singleton category trees. In general, category trees were shallow, with a maximum tree height of 5. The number of nodes within a category tree ranged from 1 to X, with the distribution of values roughly following a Poisson distribution. The number of instances per node (denoting equivalent response instances) also roughly follows a Poisson distribution, with number of instances ranging from X to Y.

@@

\subsection{Rates of Idea and Category Production}
The idea forest allows us to model the rate at which new ideas and new categories of ideas are produced, testing hypotheses 1 and 2 (i.e., rates of new ideas and categories of ideas will diminish over time, and will do so differentially per condition).

@@

We model this quantity using the following logarithmic model:
\[n_{ideas} = b_0 + b_{cond} * \log (x_{scale} * x + x_{offset})\]
where $n_{ideas}$ represents the number of new ideas, $b_0$, $x_{scale}$, and $x_{offset}$ are parameters shared between conditions that serve to position the curve, and $b_{cond}$ acts as a per-condition scaling factor. We fit the parameters to the data using a Bayesian approach [REF Kruschke], using the STAN modeling software to perform the actual fit [REF STAN]. In modeling the parameters, we use uninformative priors. The resulting posterior distributions allow us to determine whether the conditions are significantly different from one another by determining whether the highest density intervals (HDIs) of $b_{cond}$'s posterior distributions overlap or not. In our analyses, we use an HDI of 95\%.

Analysis of the HDIs of $b_{cond}$ reveals that there are no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions. [TODO: Table of fitted values and their HDIs.]

@@

We can perform a similar analysis for the number of new idea categories that arrive over time. As would be surmised from the previous graphs, we observe a similar pattern, which can also be modeled using a logarithmic model. As would also be expected, the number of new idea categories produced increases more slowly than the number of new ideas.

We see a very similar story with respect to differences in rates between conditions. The 5 and 10 conditions are virtually identical, while all other conditions are significantly different from one another. [TODO: Table of values and HDIs]

Collectively, these results largely affirm hypotheses 1 and 2: The rate of new idea and new idea categories production quickly tapers off and tends toward zero over time. Furthermore, with a few exceptions, rates differ between conditions -- new idea and idea category production is generally linked to the number of ideas requested.

Given the observed differences in production of new ideas and idea categories, it is worthwhile to better understand \emph{why} these differences exist.

@@

Early in the brainstorming process, individuals typically produce the same sets of ideas from the same sets of idea categories. It is only after the produce about 20 ideas that they start producing less generic ideas. This phenomenon can be observed by examining the o-scores of ideas and idea categories, as well as the size of the category trees. [TODO] [Mike stopped here for the night]

As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith \cite{jansson_design_1991}. An idea's o-score is $1 - p(idea)$, where $p(idea) = (number of instances of that idea)/(number of instances total)$. The o-score for category trees is similarly calculated. We will occasionally refer to the \emph{category o-score} of an idea. This simply refers to the o-score of the category tree to which that idea belongs.

We measure differences in originality between conditions by examining the idea o-score and category o-score. The distributions of these scores are summarized in Figure FIG. The left panel compares idea o-score and the right compares category o-score.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{oscores_conditions}
\end{figure}

The more ideas requested, the more original the ideas produced. As suggested by the instance quartile range in section SEC, higher number conditions produce idea in trees with fewer instances, thus the high category o-score. We also see that these conditions produce \emph{ideas} with fewer instances.

As the originality rises, however, so does the number of outliers. To understand this phenomenon, we need to more closely observe the distribution of originality scores between not just conditions, but ordinal position in a brianstorming run.

Figure FIG provides this visualization. The upper panel gives the mean idea o-score as a function of ordinal position in all 100 condition brainstorming runs. The o-score for each ordinal position is the mean of all idea o-scores for that ordinal position in a run. Following this, the plot is further smoothed by making each point the mean of a sliding window of size 10 around its ordinal position. The bottom panel is a similar plot for category-oscore. To aid in interpretation, the first, second and third quartiles are also represented on the plot. The error bars are standard error.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{idea_oscore_order_100}
    \caption{Idea o-score as a function of order (100 condition)}
    \includegraphics[width=0.9\columnwidth]{cat_oscore_order_100}
    \caption{Category o-score as a function of order (100 condition)}
\end{figure}

The only part of the originality score that falls outside the first and third quartiles is the beginning of the run. Participants in the higher numbered conditions generate more original ideas overall, but not until they have exceeded some threshold of common ideas. Figure FIG replicates figure FIG but across all conditions, and the same pattern of originality growth until an originality peak - around 20 ideas - is reached. Note that because of the additional HITs, there are far more examples of early-order ideas in the 5, 10 and 20 conditions than in the 50, 75, or 100, so this pattern is present without any dominating affect by the upper conditions.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{idea_oscore_order}
    \caption{Idea o-score as a function of order}
    \includegraphics[width=0.9\columnwidth]{cat_oscore_order}
    \caption{Category o-score as a function of order}
\end{figure}

A similar, although less severe, pattern is observed for category o-score. There is a gradual increase in category originality until the 20 instance threshold is reached, at which point variance of the scores increases rapidly and no clear pattern is observable. This is explained by the relationship between common ideas and common categories. We expect that unoriginal ideas belong to unoriginal categories, as a high instance count for an idea increases the instance count for its category tree. However, we cannot make the inverse assumption, that a high originality idea belongs to a high originality category. Once the common idea are exhausted, then, there is no relationship between idea originality and category originality.

\subsubsection{Hypothesis 3}

This interpretation supports \textbf{Hypothesis 3:} There is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition. We test this interpretation by constructing a set of the top 66 (5\%) most common ideas. If our intuition is correct, these ideas should be more likely to appear in the first 5 instances of a run than they are to appear in any of the following instances. We designate two Bernoulli variables. The first, defined by parameter $\theta_f$, represents the probability that an instance is in the common set given it is in the first five instances of a run. The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} (>5) instances in a run..

We fit these parameters using data from all runs with a total number of instances greater than 10. We chose this so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$. The resulting parameters were $\theta_f = 0.564$ (95\% HDI 0.517-0.611) and $\theta_l = 0.323$ (95\% HDI 0.303-0.342). This difference is significant, supporing the hypothesis that common ideas are over-represented in the early part of a brainstorming run, regardless of conditon.

\subsubsection{Hypothesis 6}

This threshold at 20 ideas brings us to \textbf{Hypothesis 6:} Ideas generated in the latter half of a brainstorming session are of higher originality than ideas generated in the former half. With our empirical evidence, we can instead test that ideas \emph{more than 20 instances into a brainstorming session} are more original than those in the first 20 instances, regardless of condition.

@@

\subsection{Riffing: We need to change the name of this}

At the level of the individual brainstorming run, we examine a new phenomenon related to uniqueness above. Within the SIAM model, \emph{riffing} (or \emph{clustering} as it is known in the work) occurs when a single participant activates a single image (analogous to a non-singleton category) and produces multiple ideas from that image under a fail state is reached and a transition is undergone to a new image. Unsurprisingly, we observed riffing activity in our data. An example of a single participant's riffed responses is given in Figure FIG. The second column is dark if the instance is a \emph{look-back equal} to any idea earlier in the run. The first column is dark if the instance is a \emph{source} instance, meaning it is the first in a chain of look-back equal instances. Riff chains can either be consecutive (as in the example), or separated with instances in between. The third column simply provides the text for that instance.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\columnwidth]{10_riff}
    \caption{Example of riffing in a size 10 run}
\end{figure}

The participant hits on the categories of using the device's hard drive for storage, recycling the device or its materials, and selling the device.  within these categories, they generate 2-3 ideas each. This is a fairly typical example of riffing. In larger conditions, it is common for participants to reach further back and return to an idea multiple time. Table TAB provides a summary of riffing between conditions.

@@

The higher the number of instances requested, the more participants riff on their old ideas. Similarly, upper conditions use more of their ideas as sources for riffing, and riff more times on each of those ideas. The primary surprise is that they do not riff more consecutively. The median length of a consecutive riff chain remains low across conditions. This leaves us with the image of a participant in the upper conditions searching through their old ideas for something to build upon one instance at a time.

Another outlier of note is the relatively short median distance to a previous idea in a chain for the 100 condition. Participants in the 100 condition may feel more pressure to remix, and return to earlier ideas more quickly than those in the 50 or 75 conditions.

\subsubsection{Hypothesis 4}

Riffing activity is analogous to idea production within an image as described by Nijstad and Stroebe's SIAM model \cite{nijstad_how_2006}. Thus, we should expect to confirm their hypothesis, \textbf{Hypothesis 4}, that \emph{an idea from one semantic category should more often by followed by an idea from the same category than expected by random chance}.

To test this, we modelled the probability of any category being followed by the same category with a Bernoulli distribution. Each instance not the first in its run was a trial. If the instance was in the same category tree as the previous instance, the trial was a pass. Otherwise, it was a failure. The Bernoulli $\theta$ was 0.156 (95\% HDI 0.143 - 0.170).

The most common category tree in our data set, \emph{music player}, represents only 5.65\% of instances. This is well below the 0.143 lower-bound of HDI. Thus, we accept the hypothesis that an instance follows a previous instance in the same category with greater probability than is explained by random chance.

\subsection{Completion Time}

As another aspect of how participants complete brainstorming runs, we are interested in the time spent on each response. The first, second and third quartiles for each condition are given in boxplots in figure FIG. The whiskers extend to 1.5 times the inner quartile range. Surprisingly, we see many instances taking up to 28018382ms (7.78 hours) to complete.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{time_spent_condition}
    \caption{Time spent per instance}
\end{figure}

Either our participants were exceedingly concerned with crafting their responses, or there is another influence at work. Figure FIG gives the mean time spent on an instance by order, separated by condition. Most of the highly variant response times take place early in the brainstorming run. This suggests that at least one participant accepted a brainstorming task, took an initial stab at instances, and then left the task for up to 8 hours before continuing. Beyond this, we might expect to see time spent per instance go up as participants have to dig deeper to find responses, but no such effect is consistently observed.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{time_spent_order}
    \caption{Time spent per instance}
\end{figure}

\subsubsection{Hypotheses 5}

The Nijstad and Stroede \cite{nijstad_how_2006} model also provides \textbf{Hypothesis 5}: \emph{idea generation when changing semantic categories should take longer than idea generation within categories}. An examination of the distribution of data for time spent per instance suggested a log-normal distribution, with the frequency of occurances quickly rising until a peak was reached, followed by a gradual drop-off. As with previous tests, we used the Stan package for Bayesian inference [CITE]. We fit, two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.

@@

The resulting fit models are shown in Figure FIG, overlapping the log-space histograms for observed values. The mean for the within-category condition was 9.5 in log space (13.4 seconds, 95\% HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, 95\% HDI 10.2-10.3). The variances were 1.1 (3 seconds) and 1.2 (3.3 seconds) respectively. These difference were both sigificant. Within-category instance generation took on average 12.6 seconds less than between-categories. This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds.

>>>>

Hierarchical clustering yielded an idea forest composed of 3007 instances, 1213 unique ideas, 335 category trees, and 129 non-singleton category trees. In general, category trees were shallow, with a maximum tree height of 5. The number of nodes within a category tree ranged from 1 to X, with the distribution of values roughly following a Poisson distribution. The number of instances per node (i.e., the number of equivalent idea instances) also roughly follows a Poisson distribution, with number of instances ranging from X to Y.

@@

\subsection{Rates of Idea and Category Production}
The idea forest allows us to model the rate at which new ideas and new categories of ideas are produced, testing hypotheses 1 and 2 (i.e., the notion that rates of new ideas and categories of ideas will diminish over time, and will do so as a function condition).

@@

We model this quantity using the following logarithmic model:
\[n_{ideas} = b_0 + b_{cond} * \log (x_{scale} * x + x_{offset})\]
where $n_{ideas}$ represents the number of new ideas, $b_0$, $x_{scale}$, and $x_{offset}$ are parameters shared between conditions that serve to position the curve, and $b_{cond}$ acts as a per-condition scaling factor. This model does not explicitly model variability due to different workers, a point we return to later.

We fit the parameters to the data using a Bayesian approach [REF Kruschke], using the STAN modeling software to perform the actual fit [REF STAN]. In modeling the parameters, we use uninformative priors. The resulting posterior distributions allow us to determine whether the conditions are significantly different from one another by determining whether the highest density intervals (HDIs) of $b_{cond}$'s posterior distributions overlap or not. In our analyses, we use an HDI of 95\%.

Analysis of the HDIs of $b_{cond}$ reveals no significant differences between the 5 and 10 conditions, with all other conditions significantly different from one another except for the 50 and 100 conditions. [TODO: Table of fitted values and their HDIs.] Notable in our data is the 75 condition, which rises above the 100 condition in this plot. Further analysis reveals that a single worker is responsible for that additional rise in number of new ideas. Removing that worker from the data reveals that the 75 condition is not different from either the 50 or 100 conditions.

@@

We can perform a similar analysis for the number of new idea categories that arrive over time. As would be surmised from the previous graphs, we observe a similar pattern, which can also be modeled using a logarithmic model. In the fitted model, we see that the number of new idea categories produced increases more slowly than the number of new ideas. This finding suggests that although individuals may produce new, unique ideas over time, these new ideas are within the \emph{same} categories. In other words, they are ``riffing'' -- coming up with permutations on existing ideas.

Collectively, these results largely affirm hypotheses 1 and 2: The rates of new idea and new idea categories production quickly taper off and tend toward zero over time. Furthermore, we see differences in rates between conditions. In our data, we see three clear groups of rates: the 5 and 10 conditions, the 20 condition, and the 50, 75, and 100 condition.

%Given the observed differences in production of new ideas and idea categories, it is worthwhile to better understand \emph{why} these differences exist.

@@

\subsection{Characteristics of an Idea Within a Run}
We now turn our attention to individual runs, and characterize the ideas produced throughout a run.

\subsubsection{Most Popular Ideas}
Our results suggest that individuals typically produce ideas from a common set idea categories early in the brainstorming process, then diverge and produce more original ideas. This phenomenon can be observed by examining the o-scores of ideas and idea categories in Figures X and Y, which show the distribution of o-scores for individual ideas and idea categories, as a function of response number, across all conditions.

%As our measure of originality, we use \emph{o-score}, introduced by Jansson and Smith \cite{jansson_design_1991}. An idea's o-score is $1 - p(idea)$, where $p(idea) = (number of instances of that idea)/(number of instances total)$. The o-score for category trees is similarly calculated. We will occasionally refer to the \emph{category o-score} of an idea. This simply refers to the o-score of the category tree to which that idea belongs.

% We measure differences in originality between conditions by examining the idea o-score and category o-score. The distributions of these scores are summarized in Figure FIG. The left panel compares idea o-score and the right compares category o-score.

@@

% \begin{figure}[h!]
%     \centering
%     \includegraphics[width=0.9\columnwidth]{oscores_conditions}
% \end{figure}

% The more ideas requested, the more original the ideas produced. As suggested by the instance quartile range in section SEC, higher number conditions produce idea in trees with fewer instances, thus the high category o-score. We also see that these conditions produce \emph{ideas} with fewer instances.

% As the originality rises, however, so does the number of outliers. To understand this phenomenon, we need to more closely observe the distribution of originality scores between not just conditions, but ordinal position in a brianstorming run.

% Figure FIG provides this visualization. The upper panel gives the mean idea o-score as a function of ordinal position in all 100 condition brainstorming runs. The o-score for each ordinal position is the mean of all idea o-scores for that ordinal position in a run. Following this, the plot is further smoothed by making each point the mean of a sliding window of size 10 around its ordinal position. The bottom panel is a similar plot for category-oscore. To aid in interpretation, the first, second and third quartiles are also represented on the plot. The error bars are standard error.

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=0.9\columnwidth]{idea_oscore_order_100}
%     \caption{Idea o-score as a function of order (100 condition)}
%     \includegraphics[width=0.9\columnwidth]{cat_oscore_order_100}
%     \caption{Category o-score as a function of order (100 condition)}
% \end{figure}

% The only part of the originality score that falls outside the first and third quartiles is the beginning of the run. Participants in the higher numbered conditions generate more original ideas overall, but not until they have exceeded some threshold of common ideas. Figure FIG replicates figure FIG but across all conditions, and the same pattern of originality growth until an originality peak - around 20 ideas - is reached. Note that because of the additional HITs, there are far more examples of early-order ideas in the 5, 10 and 20 conditions than in the 50, 75, or 100, so this pattern is present without any dominating affect by the upper conditions.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{idea_oscore_order}
    \caption{Idea o-score as a function of response number, all conditions}
    \includegraphics[width=0.9\columnwidth]{cat_oscore_order}
    \caption{Category o-score as a function of response number, all conditions}
\end{figure}

In both instances, we observe a steady increase in o-scores for the first 20-30 responses. After these first 20-30 responses, the o-scores essentially plateau, with no clear discernible pattern for either ideas or idea categories.

These graphs lend support for hypothesis 3, the notion that there is a set of general, common ideas that make up the first several responses of every crowd brainstorming session, regardless of condition.

To more explicitly confirm this hypothesis, we extracted the top 5\% most common ideas (66 in total). If our intuition is correct, these ideas should be more likely to appear in the first five instances of a run than they are to appear in any of the following instances. We introduce two Bernoulli random variables to model this. The first, defined by parameter $\theta_f$ represents the probability that a response instance is in the top 5\% most common ideas, given it is found in the first five instances of a run ($P(\theta_f|top)$). The second, defined by parameter $\theta_l$, is the probability that an instance is in the common pool given it is one of the \emph{last} (>5) instances in a run ($P(\theta_l|top)$).

We fit these parameters using data from all runs that have more than 10 responses. We chose this constraint so that no run could contribute evidence to $\theta_f$ without also contributing to $\theta_l$. The resulting parameters were $\theta_f = 0.564$ (95\% HDI 0.517-0.611) and $\theta_l = 0.323$ (95\% HDI 0.303-0.342). This difference is significant, supporing the hypothesis that common ideas are over-represented in the early part of a brainstorming run.

%This is explained by the relationship between common ideas and common categories. We expect that unoriginal ideas belong to unoriginal categories, as a high instance count for an idea increases the instance count for its category tree. However, we cannot make the inverse assumption, that a high originality idea belongs to a high originality category. Once the common idea are exhausted, then, there is no relationship between idea originality and category originality.

\subsubsection{Originality of Ideas}
Complementing the view of what ideas are generated when, we now consider idea and idea category o-scores as a function of response number. This analysis allows us to examine hypothesis 4, or our expectation that o-scores increase as a function of response number.

@@

\subsection{Idea Generation Patterns}
Given the clear differences in characteristics of ideas as a function of response number, we now consider idea generation patterns observable from the data. In this section, we examine these patterns in the context of hypotheses 5 and 6. Specifically, these hypotheses draw upon the SIAM model and predict that individuals will generate ideas from an idea category until they exhaust that category, at which point they switch to another category (hypothesis 5). Furthermore, SIAM predicts that this category switching will be observable as taking more time to generate an idea in a new category than in the existing category (hypothesis 6).

To determine category switching, we examine whether an idea was assigned to the same category tree as the previous, or whether it was assigned to a different category tree. We also examine the timing data collected from the web app to examine the time to generate an idea as a function of whether it is in the same category as the previous idea.

Our data show clear evidence in support of both hypotheses. As an example, below is an example of a worker riffing on a single idea.

%An example of a single participant's riffed responses is given in Figure FIG. The second column is dark if the response instance is a permutation of any idea earlier in the run. The first column is dark if the instance is a \emph{source} instance, meaning it is the source for later permutations. Permutation chains can either be consecutive (as in this example), or separated with instances in between. The third column simply provides the text for that instance.

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{10_riff}
%     \caption{Example of riffing in a size 10 run}
% \end{figure}

%The participant hits on the categories of using the device's hard drive for storage, recycling the device or its materials, and selling the device.  within these categories, they generate 2-3 ideas each. This is a fairly typical example of riffing. In larger conditions, it is common for participants to reach further back and return to an idea multiple time. Table TAB provides a summary of riffing between conditions.

@@

Examining our data, we find that the higher the number of instances requested, the more likely participants are to use old ideas as inspiration for new ideas (i.e., ``riff'' on previous ideas). However, these riff sequences are typically quite small in length, across conditions.

To explicitly test hypothesis 5, we model the probability of any idea category being followed by the same idea category with a Bernoulli random variable. Using a uniform beta prior, we find the probability of a response instance being in the same idea category as the previous to be 15.6\% (95\% HDI 14.3\% - 17.0\%). To establish a baseline to understand whether this probability is greater than chance, the most popular category tree in our data set is \emph{music player}, representing 5.65\% of all instances. Thus, the greatest probability of one instance being in the same category as the previous instance, by chance, is 5.65\%, well below the 14.3\% lower bound of the calculated HDI. Thus, we accept the hypothesis that an instance follows a previous instance in the same category with greater probability than is explained by random chance.

To test hypothesis 6, we now examine the time it takes to generate a response instance.

An examination of the distribution of data for time spent per instance suggests a log-normal distribution to the data. Using STAN, we fit two models: $t_w = \text{lognormal}(\mu_w, \sigma_w)$ for within-category idea generation and $t_b = \text{lognormal}(\mu_b, \sigma_b)$ for between-category idea generation, where the $t$ parameter in each model represents the time spent generating an instance.

@@

The resulting fit models are shown in Figure FIG, overlapping the log-space histograms for observed values. The mean for the within-category condition was 9.5 in log space (13.4 seconds, 95\% HDI 9.4-9.6), while the mean for the between-categories condition was 10.2 (27.0 seconds, 95\% HDI 10.2-10.3). The variances were 1.1 (3 seconds) and 1.2 (3.3 seconds) respectively. These difference were both sigificant. Within-category instance generation took on average 12.6 seconds less than between-categories. This is consistent with the findings of Nijstad and Stoebe, who reported differences of 6-12 seconds between conditions. Thus, we find support for hypothesis 6 and also show the applicability of the SIAM model to brainstorming in microtask marketplaces.

%Figure FIG displays the mean time spent on an instance by order, separated by condition. Most of the highly variant response times take place early in the brainstorming run. This suggests that at least one participant accepted a brainstorming task, took an initial stab at instances, and then left the task for up to 8 hours before continuing. Beyond this, we might expect to see time spent per instance go up as participants have to dig deeper to find responses, but no such effect is consistently observed.

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=0.9\columnwidth]{time_spent_condition}
%     \caption{Time spent per instance}
% \end{figure}

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=0.9\columnwidth]{time_spent_order}
%     \caption{Time spent per instance}
% \end{figure}

%The Nijstad and Stroede \cite{nijstad_how_2006} model also provides \textbf{Hypothesis 5}: \emph{idea generation when changing semantic categories should take longer than idea generation within categories}.
